% \glsresetall
\chapter{Prospect} % Main chapter title
\label{Chapter9}

\lhead{Chapter 9. \emph{Prospect}}

The time available for the research, data collection and the actual writing was limited and, therefore, this transcript is scoped to a certain degree. 
Nonetheless, the statements in this document are not final and further possibilities can be explored in that field. 
The combinational effect of the researched technologies: For instance, a WC-based micro frontend landscape, in which the resources for the components are provided by a CDN. 
Also, the WMF topic was only analyzed in the context of the UI framework Angular. 
Even though this is a valid option, the Module Federation can be used in combination with other frameworks, too. 
This is definitely a field which should be looked into. 
Especially when taking into account how it was shown that a chosen framework can affect the performance of a micro frontend. 
Another field which was not dealt with, is the development and the hosting of a CDN. 
Even though an approximate assumption was made concerning the effort connected to such a project, this is by no means empirical data. 
Therefore, since the developed landscape relies on the Unpkg.com API to request the CDN resources, it would be an interesting experiment to see if and how an own CDN could improve or optimize the performance metrics for similar landscapes.
Furthermore, it should also be inspected if a regular bundler environment with e.g. NPM could perform as good or even better as the other technologies, if a tree shaking algorithm is used \cite{treeshaking}.
Lastly, the Surge web server, used for the deployment of the landscapes, is connected to certain limitations too, namely the missing HTTP/2 server configuration. 
Thus, it would make sense to deploy those landscapes over different web servers, just to see if the changes on server-side improve the performances in the given context.

Even though the use cases covered by the prototypes are asserted to be representative, they still offer room for improvements and optimizations. Thus, when the given research is applied to a real life scenario, special conditions or requirements have to be considered when making a decision in that context. Therefore, it is mentioned in chapter \ref{Chapter7} that the gain or benefit of each respective technology is highly dependent on the individual use case. 

Nonetheless, the evidence provided in this transcript is applicable for a development team, either as a source of advice, a documentation for best practices or as a basis for design decisions. 
The goal set for this thesis can be considered to be achieved and exceeded, since it was showcased that the used technologies provide a solution for the issue of redundant libraries in micro frontend landscapes, and how they do it. 
Additionally, it was also evaluated under which circumstances those technologies are applicable under consideration of their own individual attributes. 
Based on this research, the mentioned aspects of the technologies can be explored and improved upon.




